Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-19 Manuscript/manuscript.Rmd
edits and additions to manuscript/references

Most notably is the introduction of heavily imbalanced classes. Due to the sparcity of TADs throughout the genome, most defined regions will not contain a TAD boundary. In a classification setting, when one is attempting to discriminate between regions with and without TAD boundaries, this will contribute to small imbalance ratios (defined as the ratio of minority to majority classes) [@Orriols-Puig2008]. It is commonly agreed upon in the machine learning community that imbalanced datasets adversely impact the performance of the classifiers as the learned model is biased towards the majority class to minimize the overall error rate [@lusa2010class;@chen2013influence]. Similarly, careful consideration must also be taken when deciding which performance metric(s) to present when evaluating predictive models, especially in the case of imbalanced data.

@@

In addition to employing suitable metrics for evaluating models validated on imbalanced data, there also exist several algorithmic or data level solutions. Algorithmic level solutions involve modifyng existing classifiers in order to curtail the bias induced by imbalanced classes. Many such solutions have been developed in a variety of fields related to machine learning [@estabrooks2004multiple; @chawla2003smoteboost; @chen2004using; @khoshgoftaar2007empirical]. Data level solutions mostly constitute as resampling techniques, in which random sampling is used to either replicate or removed data points in order to create a more balanced dataset.  It has been shown that quite simple resampling techniques can drastically improve model performance in the face of imbalanced data [@japkowicz2002class; @dubey2014analysis,]. For the purposes of this study we focus our attention toward data level solutions so that all results may be classifier independent, and thus, more easily comparable.

>>>>

Most notably is the introduction of heavily imbalanced classes. Due to the sparcity of TADs throughout the genome, most defined regions will not contain a TAD boundary. In a classification setting, when one is attempting to discriminate between regions with and without TAD boundaries, this will contribute to small imbalance ratios (defined as the ratio of minority to majority classes) [@Orriols-Puig2008]. It is commonly agreed upon in the machine learning community that imbalanced datasets adversely impact the performance of the classifiers as the learned model is biased towards the majority class to minimize the overall error rate [@lusa2010class;@chen2013influence;@jeni2013facing]. Similarly, careful consideration must also be taken when deciding which performance metric(s) to present when evaluating predictive models, especially in the case of imbalanced data.

@@

In addition to employing suitable metrics for evaluating models validated on imbalanced data, there also exist several algorithmic or data level solutions. Algorithmic level solutions involve modifyng existing classifiers in order to curtail the bias induced by imbalanced classes. Many such solutions have been developed in a variety of fields related to machine learning [@estabrooks2004multiple; @chawla2003smoteboost; @chen2004using; @khoshgoftaar2007empirical]. Data level solutions mostly constitute as resampling techniques, in which random sampling is used to either replicate or removed data points in order to create a more balanced dataset.  It has been shown that quite simple resampling techniques can drastically improve model performance in the face of imbalanced data [@japkowicz2002class; @dubey2014analysis]. For the purposes of this study we focus our attention toward data level solutions so that all results may be classifier independent, and thus, more easily comparable.
